---
title: "Regression Analysis 1"
author: "Eric Wang"
date: "4/16/2020"
output:
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem Number One, Part A
Draw the scatterplot of ppgdp against fertility and describe the relationship between these two variables. Is the trend linear?

ANS: Using the code below, fertility is shown to be decreasing at a faster rate and ppgdp is slowly increasing. A decreasing trend is presented. This trend seems not to be a linear trend but rather a decreasing exponential trend.
```{r UN11A}
# call for the UN11 data
library(alr4)
data(UN11)
# assigns all data in fertility category of UN11 to fertility
fertility = UN11$fertility
# assigns all data in ppgdp category of UN11 to ppgdp
ppgdp = UN11$ppgdp
# make a scatter plot with x as fertility and y as ppgdp
plot(fertility,ppgdp,xlab="Fertility",ylab="PPGDP",main="Scatter Plot")
# create a regression model naming ppgdpvsfertility
ppgdpvsfertility = lm(ppgdp~fertility)
# create a summary of the data comparison 
summary(ppgdpvsfertility)

```


## Problem Number One, Part B

Replace both variables by their natural logarithms and draw another scatterplot. Does the simple linear regression model seem plausible for a summary of this graph?

ANS: As shown below, the scatter plot presented a linear relationship between the log of fertility and the log of ppgdp. The included summary shows a adjusted residual of 0.5236. There are definitely some outliers that is affecting the fetility. Approximately, the simple linear regression model seem plausible for a summary for this graph

```{r UN11B}
# call for the UN11 data
library(alr4)
data(UN11)
# assigns all data in log(fertility) category of UN11 to fertility
fertility = log(UN11$fertility)
# assigns all data in log(ppgdp) category of UN11 to ppgdp
ppgdp = log(UN11$ppgdp)
# make a scatter plot with x as fertility and y as ppgdp
plot(fertility,ppgdp,xlab="Fertility",ylab="PPGDP",main="Scatter Plot")
summary(lm(fertility~ppgdp))

```

### Problem Number One, Part C

Without using the R function lm(),find the least-squares estimates regressing log(ppgdp) on log(fertility) and add the line in the plot in (b).

ANS: Using the code below, the computed values are fertilitybar=0.91223, ppgdpbar = 8.46379, Sfertility = 39.1982, Sppgdp = 480.47483, Sfertilityppgdp = -99.53026, beta1hat=-2.53915, and beta0hat=10.78000965.
```{r UN11C}
# call for the UN11 data
library(alr4)
data(UN11)
# assigns all data in log(fertility) category of UN11 to fertility
fertility = log(UN11$fertility)
# assigns all data in log(ppgdp) category of UN11 to ppgdp
ppgdp = log(UN11$ppgdp)
fertilitybar = mean(fertility)
## [1] 0.91223
ppgdpbar = mean(ppgdp)
## [1] 8.46379
# sum of squres, using the formula
Sfertility = sum((fertility-fertilitybar)^2)
## [1] 39.1982
Sppgdp = sum((ppgdp-ppgdpbar)^2)
## [1] 480.47483
Sfertilityppgdp = sum((fertility-fertilitybar)*(ppgdp-ppgdpbar))
## [1] -99.53026
# estimate of the value of the slope with the formula from lecture
beta1hat = Sfertilityppgdp/Sfertility
## [1] -2.53915
# estimate if the value of the intercept with the formula provided in lecture
beta0hat = ppgdpbar-beta1hat*fertilitybar
## [1] 10.78000965
# fitted values calculation
ppgdphat = beta0hat+beta1hat*fertility
# Draw a scatter plot using plot(), including the fitted line
plot(fertility,ppgdp,xlab="lcavol",ylab="lpsa",main="lpsa vs lcavol",abline(beta0hat,beta1hat,col= 'red'))
```

#### Problem Number Two, Part A

Draw a scatterplot. Does a simple linear regression model seem reasonable?

ANS: A simple linear regression model seem reasonable because as lcavol increases, lpsa increases as well. The graph shows a linear positive relationship between lpsa and lcavol.
```{r prostateA}
# retrieve the data in prostate
library(faraway)
data(prostate)
# assign data in the lcavol category to lacavol
x = prostate$lcavol
# assign data in the lpsa category to lpsa
y = prostate$lpsa
# Draw a scatter plot using plot()
plot(x,y,xlab="lcavol",ylab="lpsa",main="lpsa vs lcavol")
```

##### Problem Number Two, Part B

Without using the R function lm(), compute the values x;  Y ; SXX; SY Y and SXY . Compute the ordinary least squares estimates of the intercept and slope for the simple linear regression model, and draw the fitted line on your plot from part (a).

ANS: Using the code below, the computed values are xbar=1.350009579, ybar = 2.4783870103, Sx = 133.359034, Sy = 127.9175839930, Sxy = 95.9278, beta1hat=0.71932, and beta0hat=1.50729793.
```{r prostateB}
# retrieve the data in prostate
library(faraway)
data(prostate)
# assign data in the lcavol category to lacavol
x = prostate$lcavol
y = prostate$lpsa
# means of the sample
xbar = mean(x)
## [1] 1.350009579
ybar = mean(y)
## [1] 2.4783870103
# sum of squres, using the formula
Sx = sum((x-xbar)^2)
## [1] 133.359034
Sy = sum((y-ybar)^2)
## [1] 127.9175839930
Sxy = sum((x-xbar)*(y-ybar))
## [1] 95.9278
# estimate of the value of the slope with the formula from lecture
beta1hat = Sxy/Sx
## [1] 0.71932
# estimate of the value of the intercept with the formula provided in lecture
beta0hat = ybar-beta1hat*xbar
## [1] 1.50729793
# fitted values calculation
yhat = beta0hat+beta1hat*x
# Draw a scatter plot using plot(), including the fitted line
plot(x,y,xlab="lcavol",ylab="lpsa",main="lpsa vs lcavol", abline(beta0hat,beta1hat,col= 'red'))
```

###### Problem Number Two, Part C

Obtain the estimate of sigma squared and find the estimated standard errors of bhat0 and bhat1.

ANS: Using the code below, the computed values are sse = 58.9147574, mse = 0.62015534, seb1 = 0.06819, and seb0 = 0.1219368

```{r prostateC}
#get the number of observations
n = length(x)
## [1] 97L
# apply the formula for sum of square error
sse = Sy-beta1hat*Sxy
## [1] 58.9147574
# apply the mean square error formula
mse = sse/(n-2)
## [1] 0.62015534
# standard errors estimation
seb1 = sqrt(mse/Sx)
## [1] 0.06819
seb0 = sqrt(mse*sum(x^2)/(n*Sx))
## [1] 0.1219368

```

###### Problem Number Two, Part D

Find the estimated covariance between b0 and b1, i.e. Cov b0; b1 b:

ANS: Using the code below, the estimated computed value is cov = -0.006277907287

```{r prostateD}
# estimated cov between bhat0 and bhat1
cov = -mse*xbar/Sx
## [1] -0.006277907287

```

###### Problem Number Two, Part E

Carry out t-tests for the two null hypotheses 0 = 0 and 1 = 0, reporting the value of the test statistic and a p-value in each case.

ANS: Using the code below, the computed values are t_statb0 = 12.3613022211, p_valb0 = 0.000125, t_statb1 = 10.5483175586397, and p_valb1 = 1.11861577347983e-17

```{r prostateE}
# beta 0 test statistics
t_statb0 = beta0hat/seb0
## [1] 12.3613022211

# beta 0 p- value, using the formula
p_valb0 = pt(abs(t_statb0),df=n-2,lower.tail=FALSE)+ pt(-abs(t_statb0),df=n-2,lower.tail=TRUE)
## [1] 1.72223426077146e-21

# beta 1 test statistics
t_statb1 = beta1hat/seb1
## [1] 10.5483175586397
# beta 1 p- value, using the formula
p_valb1 = pt(abs(t_statb1),df=n-2,lower.tail=FALSE)+ pt(-abs(t_statb1),df=n-2,lower.tail=TRUE)
## [1] 1.11861577347981e-17
```

###### Problem Number Two, Part F

Use the R function lm() to fit the regressions of lpsa on lcavol.

```{r prostateF}
# using the built in function lm to fir the regressions of lpsa on lcavol, given that the data are from prostate
lpsavslcavol = lm(lpsa~lcavol, data=prostate)
# create a summary report
summary(lpsavslcavol)


```










