---
title: "Regression Analysis 4"
author: "Eric Wang"
date: "5/13/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ## Problem One, Part A

Examine the scatterplot matrix for (fertility, log(ppgdp), pctUrban), and comment on the marginal relationships.

ANS: From the scatter matrix for (fertility, log(ppgdp), pctUrban), we can say that Fertility and log(ppgdp) has a negative correlation, Fertility and pctUrban has a negative correlation, and log(ppgdp) and pctUrban has a positive correlation

```{r cars}
library(alr4) 
data(UN11)
# find the scatter matrix
pairs(fertility ~ log(ppgdp)+pctUrban,data=UN11) 

```

## Problem One, Part B

Fit the two simple regressions fertility  log(ppgdp) and fertility  pctUrban, and verify that the slopes are signicantly dierent from zero at any conventional level of significance.

ANS: The slope of log(ppgdp) is siginificantly different from zero at any conventional level of significance as the p value is relatively lower than any alpha value.

```{r 1}
# fit the model
fit0=lm(fertility~log(ppgdp),data=UN11)
fit1=lm(fertility ~ pctUrban, data= UN11)
summary(fit0)
summary(fit1)


```
## Problem One, Part C

Obtain the added-variable plots for both predictors. Based on the added-variable plots, does log(ppgdp) seem to be useful after adjusting for pcturban, and similarly, does pcturban seem to be useful after adjusting for log(ppgdp)?

ANS: From the avPlot, log(ppgdp) seems to be useful as it has a steep slope. On the otherhand, pctUrban also seems to not be as useful as it relatively neutral against fertility.

```{r 2}
# added variable plot
av = avPlots(lm(fertility~log(ppgdp)+pctUrban,data=UN11))

```
## Problem Two, Part A

Examine the scatterplots for all variables, and comment on the marginal relationships.

ANS: From the scatter matrix, we can say that BSAAM and OPBPC has a postitive correlation, BSAAM and OPRC has a postive correlation, and BSAAM OPSLAKE has a negative correlation.


```{r 3}
library(alr4)
data("water")
# find the scatter matrix
pairs(BSAAM~OPBPC + OPRC + OPSLAKE, data=water)



```
## Problem Two, Part B

Fit the three simple regressions for BSAAM  OPBPC, BSAAM  OPRC, and BSAAM  OPSLAKE and verify that the slope coecients are signicantly dierent from 0 at any conventional level of signicance.

ANS: The slope of OPRC, OPSLAKE, and OPBPC is siginificantly different from zero at any conventional level of significance as the p value is relatively lower than any alpha value.

```{r 4}
# fit the model
fit2=lm(BSAAM~OPBPC, data = water)
fit3=lm(BSAAM~OPRC, data = water)
fit4=lm(BSAAM~OPSLAKE, data = water)

summary(fit2)
summary(fit3)
summary(fit4)
```
## Problem Two, Part C

Obtain the added-variable plots for the model. Report your findings.

ANS: For the OPBPC predictor avplots, the predictor OPBPC contains no additional useful information in predicting BSAAM
     For the OPRC predictor avplots, the predictor OPRC contains significant information in predicting BSAAM
     For the OPSLAKE predictor avplots, the predictor OPSLAKE contains significant information in predicting BSAAM

```{r 5}
# added variable plot
fit5=lm(BSAAM~OPBPC+OPRC+OPSLAKE, data = water)
avPlots(fit5)

```
## Problem Two, Part D

Get the regression summary for the regression of BSAAM on these three regressors. Include OPBPC, OPRC, and OPSLAKE sequentially. Explain the values on the \Pr(> jtj)" column of your output, with given  = :05.

ANS: IF the p-value is greater than the alpha value, then we fail to reject Ho. If the p-value is less than the alpha value, then we reject Ho.
     For the "(intercept)", since the p-value is 1.1e-0.7 and the given alpha value is 0.05. The p-value is less than the alpha value. Thus, we reject Ho.
     For the OPBPC, since the p-value is 0.93599 and the given alpha value is 0.05. The p-value is greater than the alpha value. Thus, we fail to reject Ho.
     For the OPRC, since the p-value is 0.00633 and the given alpha value is 0.05. The p-value is less than the alpha value. Thus, we reject Ho.
     For the OPSLAKE, since the p-value is 0.00410 and the given alpha value is 0.05. The p-value is is less than the alpha value. Thus, we reject Ho.

```{r 6}
summary(fit5)

```
## Problem Two, Part E

Interpret the appropriate R2 value in (d).

ANS: The Adjusted R-squared is 0.8941 which is close to 1.This means that 89.41% of variablity in the response is explained by a MLR model with all the predictors collectively.. This also means that there is a strong linear relationship among all predictors.

## Problem Two, Part F

Use a built-in function, anova(), in R, to conduct a global F-test of whether or not
1 = 2 = 3 = 0 given  = :05. State the alternative hypothesis, p-value, decision rule,
and conclusion.

ANS: Null hypothesis: BetaOPBPC = BetaOPRC = BetaOPSLAKE = 0
     Alternative Hypothesis: for some Beta j is not equal to zero, j=1,2,3 (or at least one of Beta j is not equal to zero).
     Given the alpha value is 0.05, the p-value of the Gobal F test is less than 2.2e-16 (is close to zero). And all p-values are less than the alpha value of 0.05. 
     Thus, we reject Ho as a result.
     This means that for some or atleast one of Beta j is not equal to zero

```{r 7}
# anova table of fit5
anova(fit5)


```
## Problem Three, Part A

Fit a model with total sat score as the response and expend, ratio and salary as predictors sequentially.
1. Test the hypothesis that 3 = 0; i.e. salary = 0; at  = 0:05: What is your conclusion for this test?

   ANS: Since the p-value against salary is 0.0667 is greater than the alpha value of 0.05, we fail to reject the null hypothesis. 
   
2. Test the hypothesis that 1 = 2 = 3 = 0 i.e. expend = ratio = salary = 0; at alpha = 0:05: What is your conclusion for this test?

   ANS: Using the global F test, the p- value of the global F test is 0.01209 which is less than the 0.05 alpha value. THis means that we reject Ho and Ho is not true. 
   
3. Do any of these predictors have an eect on the response?

   ANS: At least one of them is different from zero and at least one of them have an effect on the response.
```{r 8}
library(faraway)
data("sat")
# fit the model
fit6 = lm(total ~ expend + ratio + salary, data = sat)
summary(fit6)

```
## Problem Three, Part B

Now add takers to the model. Test the hypothesis that takers = 0; at  = 0:05: What is your conclusion? To conrm your conclusion, draw the added-variable plots for the model.
Report your findings.

ANS: We reject Ho as well since p-value of the Global F test is less than the 0.05 alpha value.
```{r 9}
library(faraway)
data("sat")
# fit the model
fit6b = lm(total ~ expend + ratio + salary + takers, data = sat)
summary(fit6b)
# added variable plot
av6b = avPlots(fit6b)

```
