---
title: "Regression Analysis 2"
author: "Eric Wang"
date: "4/22/2020"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PROBLEM NUMBER ONE, Part A

Do the regression of length on age, and report the estimates, their standard errors and the estimate of variance, i.e. report ^ 0; ^ 1; SE( ^ 0), SE( ^ 1) and MSE.

ANS: b1hat = 30.323889, b0hat = 65.52716, SEb1 = 0.6877, SEb0 = 3.197, MSE = 820.5847
```{r alr4 wblake a}
library(alr4)
data(wblake)
names(wblake)

x = wblake$Age
y = wblake$Length
xbar = mean(x)
## [1] 4.2027
ybar = mean(y)
## [1] 192.970387
n = length(y)
## [1] 439L
Sxy = sum((x-xbar)*(y-ybar))
## [1] 52610.6355
Sxx = sum((x-xbar)^2)
## [1] 1734.9567
Syy = sum((y-ybar)^2)
## [1] 1953954.61503

# regression estimates
b1hat = Sxy/Sxx 
## [1] 30.323889
b0hat = ybar-b1hat*xbar
## [1] 65.52716
# regression of length on age
yhat = b0hat+b1hat*x 
SSE = sum((y-yhat)^2)
## [1] 358595.52635
SSR = sum((yhat-ybar)^2)
## [1] 1595359.08868
SST = SSR + SSE
## [1] 1953954.6150
# coefficient of determination
RSq = SSR/SST
## [1] 0.8165
# variance estimation
MSE = SSE/(n-2)
## [1] 820.5847
# slope b1 standard error
SEb1 = sqrt(MSE/Sxx)
## [1] 0.6877
# intercept b0 standard error
SEb0 = sqrt(MSE*((1/n)+(xbar^2/Sxx)))
## [1] 3.197

```

## PROBLEM NUMBER ONE, Part B

Is it meaningful to interpret b0? Why or why not?

ANS: No, it is not meaningful to interpret b0hat because the provided age of data(x value) does not include 0, which essentially means that x never equals 0. so b0hat has virtually no meaning.

```{r alr4 wblake b}
```

## PROBLEM NUMBER ONE, Part C

Interpret b1

ANS: the interpretation of b1hat is that for every increase in small mouth bass' age (x value), the small mouth bass length is expected to increase by 30.23389

```{r alr4 wblake C}
```

## PROBLEM NUMBER ONE, Part D

Obtain a 90% condence interval for 1 from the data. Interpret.

ANS: we are 90% confident that for an unit of increase in small mouth bass age (x value), the small mouth bass will grow in length between 29.19051mm and 31.45727mm


```{r alr4 wblake D}
# 90% confidence interval for b1
a = 1-0.90
## [1] 0.1
# Critical value t
t = qt(1-a/2,n-2)
## [1] 1.648
# Margin of error
ME = t*SEb1
## [1] 1.1336
# 90% confidence interval for b1
# combine ME and -ME into a list and subtract
b1hat-c(ME,-ME)
## [1] 29.19027  31.45751
```

## PROBLEM NUMBER ONE, Part E

What is the expected length for all small mouth bass if the age of the bass is 1 at capture? And obtain a 90% condence interval for the mean length. Interpret this interval.

ANS: We are 90% confident that the expeted length of all small mouth bass at the age of 1 is between the interval of [91.57782,100.12429]
```{r alr4 wblake E}
# Age of the bass
x0 = 1
# finding the expected y0
y0 = b0hat+b1hat*x0 
## [1] 95.85105
# alpha
a = 1-0.9
## [1] 0.1
# t critical value
t = qt(1-a/2,n-2)
## [1] 1.6483
ME = t * sqrt(MSE*((1/n)+((x0-xbar)^2)/Sxx))
## [1] 4.273234
# combine ME and -ME into a list and subtract
y0 - c(ME,-ME)
## [1]91.57782  100.12429
```

## PROBLEM NUMBER ONE, Part F

Obtain a prediction and a 90% prediction interval for a small mouth bass at age 1. Interpret this interval.

ANS: When the small mouth bass is at age 1, the predicted length is 95.851mm. we are 90% confident that for a value increase in small mouth bass age (x value), the small mouth bass will have a length between 48.44976mm and 143.25234mm
```{r alr4 wblake F}
# 90% prediction for a small mouth bass at age 1
x0 = 1
# the value of predicted y at x = 1
pred_y = b0hat+b1hat*x0
## [1] 95.851
ME = t*sqrt(MSE)*sqrt(1+(1/n)+((x0-xbar)^2)/Sxx)
## [1] 47.411
# combine ME and -ME into a list and subtract
pred_y - c(ME,-ME)
## [1] 48.43975  143.26235
```

## PROBLEM NUMBER Two, Part A

Fit the regression of the response on the predictor. Draw a scatterplot of the data and add your tted regression line.

```{r alr4 Heights A}
library(alr4)
data("Heights")
x = Heights$mheight
y = Heights$dheight
# regression model 
model2=lm(y~x,data = Heights)
plot(x,y, xlab = "mheight", ylab = "dheight")
abline(model2, col = "red")


```

## PROBLEM NUMBER Two, Part B

Report the estimates, their standard errors, the value of the coecient of determination, and the estimate of variance. Interpret the coecient of determination

ANS: b1hat = 0.541747, b0hat = 29.9174, SEb1 = 0.02596, SEb0 = 1.622, RSq = 0.240795677, MSE = 5.136

    Interpretation of coefficient of determination:
        The coefficient of determination is 0.2408. This means that 24.08% of the variation in y is explained by the variation in x. (100-24.08)% is not explained.

```{r alr4 Heights B}
summary(model2)
xbar = mean(x)
## [1] 62.4528
ybar = mean(y)
## [1] 63.751
n = length(y)
## [1] 1375L
Sxy = sum((x-xbar)*(y-ybar))
## [1] 4128.40344
Sxx = sum((x-xbar)^2)
## [1] 7620.90672
Syy = sum((y-ybar)^2)
## [1] 9288.61597

# regression estimates
b1hat = Sxy/Sxx 
## [1] 0.541747
b0hat = ybar-b1hat*xbar
## [1] 29.9174
# regression of length on age
yhat = b0hat+b1hat*x 
SSE = sum((y-yhat)^2)
## [1] 7051.957
SSR = sum((yhat-ybar)^2)
## [1] 2236.658575
SST = SSR + SSE
## [1] 9288.61597
# coefficient of determination
RSq = SSR/SST
## [1] 0.240795677
# variance estimation
MSE = SSE/(n-2)
## [1] 5.136
# slope b1 standard error
SEb1 = sqrt(MSE/Sxx)
## [1] 0.02596
# intercept b0 standard error
SEb0 = sqrt(MSE*((1/n)+(xbar^2/Sxx)))
## [1] 1.622
```

## PROBLEM NUMBER Two, Part C

Obtain a estimated mean of daughters' heights, and a 99% condence interval for daughters whose mother is 60 inches tall. Interpret this interval.

ANS:
```{r alr4 Heights C}
# mothers height
x0 = 60
# daughters height estimation
y0 = b0hat+b1hat*x0 
## [1] 62.422257
a = 1-0.99
## [1] 0.01
# t critical value
t = qt(1-(a/2),n-2)
## [1] 2.579
ME = t * sqrt(MSE*((1/n)+((x0-xbar)^2)/Sxx))
## [1] 0.22766
# combine ME and -ME into a list and subtract
y0 - c(ME,-ME)
```

## PROBLEM NUMBER Two, Part D

Obtain a predicted value and 99% prediction interval for a daughter whose mother is 60 inches tall. Interpret this interval.

ANS: We are 99% confident that if mheight is 60 inches tall, the dheight will be somewhere between 56.57 and 68.27
```{r alr4 Heights D}
# 909% confidence interval for b1
a = 1-0.99
## [1] 0.01
n = length(y)
## [1] 1375L
# Critical value t
tCV = qt(1-a/2,n-2)
## [1] 2.5794
ME = tCV*SEb1
## [1] 5.850
# 99% prediction for a daughter whose mother is 60 inches tall
x0 = 60
# the value of predicted y at x = 60
pred_y = b0hat+b1hat*x0
## [1] 62.422257
ME = tCV*sqrt(MSE)*sqrt(1+(1/n)+((x0-xbar)^2)/Sxx)
## [1] 5.85
# combine ME and -ME into a list and subtract
pred_y - c(ME,-ME)
## [1] 56.57207  68.27245



```

## PROBLEM NUMBER Two, Part E

Compute the (Pearson) correlation coecient rXY by using R2 value. What does the
value of rXY imply about the relationship between dheight and mheight?

ANS: The correlation coefficient between the daughter and the mother is 0.4907094.This means that there exist a postive moderately lineaer correlation between the height of the mother and the daighter
```{r alr4 Heights E}
cor(Heights$mheight, Heights$dheight)
## [1] 0.4907094
```

## PROBLEM NUMBER Three, Part A

Use the lm function in R to t the regression of the response on the predictor. Draw ascatterplot of the data and add your tted regression line.

```{r alr4 ftcollinssnow A}
library(alr4)
data("ftcollinssnow")
early = ftcollinssnow$Early
late = ftcollinssnow$Late
model3 = lm (late ~ early, data = ftcollinssnow)
summary (model3)
plot(early,late, xlab = "Early", ylab = "Late")
abline(model3, col = "red")
```

## PROBLEM NUMBER Three, Part B

Test the null hypothesis that the slope is 0 against a two-sided alternative at  = 0.01, and interpret your ndings.

ANS: From the summary table, we see that the p-value = 0.1239, which is greater than the alpha value 0.01 , we fail to reject the null hypothesis. Which could mean that there is no possible linear relationship between late and early.
```{r alr4 ftcollinssnow B}


```

## PROBLEM NUMBER Three, Part C

What percentage of the variability in Late snowfall is explained by Early snowfall?

ANS: As we can see from the summary table from part a, the multiple r-squared is 0.02581. Thus, the percent of the variablility in late snowfall explained by early snowfall is 0.02581.

```{r alr4 ftcollinssnow C}


```

## PROBLEM NUMBER Four, Part A

Use R to produce an ANOVA table for this regression t. Test the null hypothesis that the slope is 0 against a two-sided alternative at the default , and interpret your findings.

ANS: As shown on the table below, the f statistics is 111.3. And also the p-value is less than the value of alpha. As a result, we reject the null hypothesis and slope coefficient != zero.

```{r faraway prostate A}
library(faraway)
data("prostate")
lcavol = prostate$lcavol
lpsa =  prostate$lpsa
model4 = lm(lpsa~lcavol)
summary(model4)
anova(model4)


```

## PROBLEM NUMBER Four, Part B

(b) Using the ANOVA table from part (a), calculate the coefficient of determination. Interpret your finding.
 ANS: SSR = 69.003
      SSE = 58.915
      SST = SSR + SSE = 127.918
      Coefficient of determination = 1-(58.915/127.918) = 0.5394
        This means that 53.94% of variability in lpsa is explained by the model 
        and (100-53.94)% = 46.06% of variability in lpsa are not explained by the         model
